---
title: "Assignment 1"
author: "Eunhee Lee"
date: "April 25, 2017"
output:
  pdf_document: default
  html_document: default
---


## Average temperature by month from 1942 to 2016

```{r eval=TRUE, echo=TRUE, warning=FALSE}
clim = read.table("clim.txt", header=TRUE)
Aggdata <- aggregate(clim, by=list(clim$month, clim$year), FUN="mean")
Aggdata$Tmean <- (Aggdata$tmax+Aggdata$tmin)/2
boxplot(Aggdata$Tmean ~ Aggdata$month, ylab="Temperature(C)", xlab="month", col="orange")
```



## Precipitation by month from 1942 to 2016
```{r eval=TRUE, echo=TRUE, warning=FALSE}
clim2 <- clim[,2:7]
Aggrain <- aggregate(clim2, by=list(clim2$month, clim2$year), FUN="sum")
boxplot(Aggrain$rain~Aggrain$Group.1, ylab=" Precipitation(mm)", xlab="month", col="blue")
```

## The wettest year
```{r eval=TRUE, echo=TRUE, warning=FALSE}
# Aggregate precipitation by year 
Aggdata2 <- aggregate(clim2, by=list(clim2$year), FUN="sum")

# what is the wettest year (total precipitation per year)
Wetyear=which.max(Aggdata2$rain)
Aggdata2[Wetyear,c("Group.1","rain")]
Aggdata2$year= as.integer(Aggdata2$Group.1)
Aggdata2$rain = as.integer(Aggdata2$rain)
Aggdata2[Wetyear,c("year","rain")]
```


## The driest year
```{r eval=TRUE, echo=TRUE, warning=FALSE}
# what is the driest year (total precipitation per year)
Dryyear=which.min(Aggdata2$rain)
Aggdata2[Dryyear,c("Group.1","rain")]
Aggdata2[Dryyear,c("year","rain")]

```

## The picture of a dry year

<div align="left">
<img src = "drought.jpg" height=300, width=400>
</div>

## The picture of a wet year

<div align="left">
<img src = "flood.jpg" height=300, width=400>
</div>


## Addition of season column

```{r eval=TRUE, echo=TRUE, warning=FALSE}
clim2$Season <- ifelse(clim2$month <= 5 & clim2$month >= 3, 1, ifelse(clim2$month <= 8, 2,(ifelse(clim2$month <= 11, 3, 4))))
```

## The wettest season (mean season precipitation)
```{r eval=TRUE, echo=TRUE, warning=FALSE}
Aggdata4 <- aggregate(clim2, by=list(clim2$Season, clim2$year), FUN="sum")
Seasonagg <- Aggdata4[,c(1,2,5)]
colnames(Seasonagg) <- c("Season", "year", "rain")
Aggdata3 <-aggregate(Seasonagg, by=list(Seasonagg$Season), FUN="mean")
Wetseason = which.max(Aggdata3$rain)
Aggdata3[Wetseason,c("Group.1","rain")]
Aggdata3$rain = as.integer(Aggdata3$rain)
Aggdata3[Wetseason,c("Season","rain")]

```


## The driest season (mean season precipitation)
```{r eval=TRUE, echo=TRUE, warning=FALSE}
Dryseason = which.min(Aggdata3$rain)
Aggdata3[Dryseason,c("Group.1","rain")]
Aggdata3$rain = as.integer(Aggdata3$rain)
Aggdata3[Dryseason,c("Season","rain")]
```

## Relationship between winter precipitation and summer temperature 
```{r eval=TRUE, echo=TRUE, warning=FALSE}

Aggdata5 <- aggregate(clim2, by=list(clim2$Season, clim2$year), FUN="mean")
Aggdata5$Tmean <- (Aggdata5$tmax + Aggdata5$tmin)/2
Summertem <- subset(Aggdata5, Season == 2)
Winterrain <- subset(Seasonagg, Season == 4)
Reltemrain <- merge(Summertem, Winterrain,by="year")

library(ggplot2)
ggplot(Reltemrain, aes(Tmean,rain.y))+
  geom_point(size=2)+ # creat scatter plot, and dot as blank circles
  geom_smooth(method="lm", se=FALSE)+ # creat a trend line
  xlab("Average Summer Temperature(C)")+
  ylab("Winter Precipitation(mm)") +
  theme_bw() +
  theme(text=element_text(family ="Times", size=14), panel.grid.major = element_blank(),panel.grid.minor = element_blank())

lmtemrain <- lm(rain.y ~ Tmean, data=Reltemrain)
summary(lmtemrain) 


```

As the trend line in the graph shows, the linear regression revealed that the winter precipitation decreases slightly as the average summer temperature increases. However, the precipitation only declines by 1mm during winter when the average temperature rises by 1&deg;C, showing a nearly flat line. This result indicates that the winter precipitation may not have a significant relationship with the summer temperature.  



